Unbalanced Mallows Models for Optimizing Expensive Black-Box Permutation Problems

### 2021-02-04 paper/main.tex

\subsubsection{UMM}
%
UMM uses a budget of evaluations ($\FEmax$) of  $\minit=10$ solutions sampled uniformly.
The other two parameters of UMM are $r_1$ and $r_2$ that determine the learning rate $\rho$ in uBorda, and consequently, determine the
balance between exploration and exploitation. In the experimental section we
will study their impact.

In all
experiments, we consider a maximum budget of $\FEmax=400$ evaluations of the
actual objective function. In a white-box context, state-of-the-art algorithms
for the LOP and PFSP typically evaluate thousands of solutions,
thus, the budget considered here for the black-box context is extremely
limited.

\subsubsection{CEGO} \EKHINE{tenemos una pagina entera dedicada a CEGO :s}\MANUEL{Sugiere qué quitar usando el commando \textbackslash{}ekhine}
Combinatorial Efficient Global Optimization (CEGO)~\citep{ZaeStoFriFisNauBar2014} is an extension of the well-known EGO
method~\citep{JonSchWel98go} to unconstrained black-box combinatorial
optimization problems. In EGO, Gaussian process models are used as a surrogate
of the landscape of the expensive original problem. An optimization method
searches for solutions in the surrogate model by optimizing the expected
improvement criterion, which balances the expected mean and variance of the
chosen solution. Once a solution is chosen, it is evaluated on the actual
fitness function and the result is used to update the surrogate-model,
hopefully increasing its predictive power.

@@

\subsubsection{Other settings}
%
For simplicity, we use Kendall's-$\tau$
distance~(Eq.~\ref{eq:kendall_decomp}) in all experiments.
%Nevertheless,
%\citet{ZaeStoFriFisNauBar2014} points out that CEGO performs best on the QAP
%when using Hamming distance and the same may apply to UMM.
We plan to extend UMM to other distance measures in the future.

\subsubsection{Benchmark problems}
%
Experiments with real-world expensive black-box problems would be computationally infeasible. Instead, we consider classical combinatorial optimization problems as black-box optimization problems. For consistency with the rest of the paper, a permutation $\sigma$ denotes a ranking also in this section. Recall that the ordering associated to ranking $\sigma$ is given by its inverse, $\sigma^{-1}$. Thus, we will use ordering $\sigma^{-1}$ to formulate the problems. Later in the experimental section we discuss the effects of not performing this inversion.

@@

\begin{figure*}[tb]
\small
\centering%
\begin{minipage}{0.49\linewidth}
  \texttt{N-t59b11xx} (LOP)\\[-0.5ex]
  \includegraphics[width=\textwidth]{../img/fitness_lop_IO_N-t59b11xx}
\end{minipage}
\begin{minipage}{0.49\linewidth}
  \texttt{N-sgb75-01} (LOP)\\[-0.5ex]
  \includegraphics[width=\textwidth]{../img/fitness_lop_SGB_N-sgb75_02}
\end{minipage}
\\
\begin{minipage}{0.49\linewidth}
  \texttt{N-p40-01} (LOP)\\[-0.5ex]
  \includegraphics[width=\textwidth]{../img/fitness_lop_RandB_N-p40-01}
\end{minipage}
\begin{minipage}{0.49\linewidth}
  \texttt{N-p50-01} (LOP)\\[-0.5ex]
  \includegraphics[width=\textwidth]{../img/fitness_lop_RandB_N-p50-01}
\end{minipage}
\\
\begin{minipage}{0.49\linewidth}
  \texttt{reC19} (PFSP)\\[-0.5ex]
  \includegraphics[width=\textwidth]{../img/fitness_pfsp_rec05_txt}
\end{minipage}
\begin{minipage}{0.49\linewidth}
  \texttt{reC31} (PFSP)\\[-0.5ex]
  \includegraphics[width=\textwidth]{../img/fitness_pfsp_rec13_txt}
\end{minipage}
\\
\begin{minipage}{0.49\linewidth}
  \texttt{reC19} (PFSP)\\[-0.5ex]
  \includegraphics[width=\textwidth]{../img/fitness_pfsp_rec19_txt}
\end{minipage}
\begin{minipage}{0.49\linewidth}
  \texttt{reC31} (PFSP)\\[-0.5ex]
  \includegraphics[width=\textwidth]{../img/fitness_pfsp_rec31_txt}
\end{minipage}
\label{fig:results}
\caption{Mean fitness  (and standard deviation)  of each solution evaluated by each algorithm on various instances.\label{fig:results}}
\end{figure*}

>>>>

\myparagraph{UMM}
%
UMM uses a budget of evaluations ($\FEmax$) of  $\minit=10$ solutions sampled uniformly.
The other two parameters of UMM are $r_1$ and $r_2$ that determine the learning rate $\rho$ in uBorda, and consequently, determine the
balance between exploration and exploitation. In the experimental section we
will study their impact.

We consider 3 different values for the maximum budget of $\FEmax=100, 200,400$ evaluations of the
actual objective function. In a white-box context, state-of-the-art algorithms
for the LOP and PFSP typically evaluate thousands of solutions,
thus, the budget considered here for the black-box context is extremely
limited.

Combinatorial Efficient Global Optimization (CEGO)~\citep{ZaeStoFriFisNauBar2014} is an extension of the well-known EGO
method~\citep{JonSchWel98go} to unconstrained black-box combinatorial
optimization problems. In EGO, Gaussian process models are used as a surrogate
of the landscape of the expensive original problem. An optimization method
searches for solutions in the surrogate model by optimizing the expected
improvement criterion, which balances the expected mean and variance of the
chosen solution. Once a solution is chosen, it is evaluated on the actual
fitness function and the result is used to update the surrogate-model,
hopefully increasing its predictive power.

@@

\myparagraph{Other settings}
%
For simplicity, we use Kendall's-$\tau$
distance~(Eq.~\ref{eq:kendall_decomp}) in all experiments.
%Nevertheless,
%\citet{ZaeStoFriFisNauBar2014} points out that CEGO performs best on the QAP
%when using Hamming distance and the same may apply to UMM.
We plan to extend UMM to other distance measures in the future.

\myparagraph{Benchmark problems}
%
Experiments with real-world expensive black-box problems would be computationally infeasible. Instead, we consider classical combinatorial optimization problems as black-box optimization problems. For consistency with the rest of the paper, a permutation $\sigma$ denotes a ranking also in this section. Recall that the ordering associated to ranking $\sigma$ is given by its inverse, $\sigma^{-1}$. Thus, we will use ordering $\sigma^{-1}$ to formulate the problems. Later in the experimental section we discuss the effects of not performing this inversion.

@@

\begin{figure*}[tb]
\small
\centering%
\begin{minipage}{0.49\linewidth}
  \texttt{N-p40-01} (LOP)\\[-0.5ex]
  \includegraphics[width=\textwidth]{../img/fitness_lop_RandB_N-p40-01}
\end{minipage}
\begin{minipage}{0.49\linewidth}
  \texttt{N-p50-01} (LOP)\\[-0.5ex]
  \includegraphics[width=\textwidth]{../img/fitness_lop_RandB_N-p50-01}
\end{minipage}
\\
\begin{minipage}{0.49\linewidth}
  \texttt{N-t59b11xx} (LOP)\\[-0.5ex]
  \includegraphics[width=\textwidth]{../img/fitness_lop_IO_N-t59b11xx}
\end{minipage}
\begin{minipage}{0.49\linewidth}
  \texttt{N-sgb75-01} (LOP)\\[-0.5ex]
  \includegraphics[width=\textwidth]{../img/fitness_lop_SGB_N-sgb75_02}
\end{minipage}
\\
\begin{minipage}{0.49\linewidth}
  \texttt{reC19} (PFSP)\\[-0.5ex]
  \includegraphics[width=\textwidth]{../img/fitness_pfsp_rec05_txt}
\end{minipage}
\begin{minipage}{0.49\linewidth}
  \texttt{reC31} (PFSP)\\[-0.5ex]
  \includegraphics[width=\textwidth]{../img/fitness_pfsp_rec13_txt}
\end{minipage}
\\
\begin{minipage}{0.49\linewidth}
  \texttt{reC19} (PFSP)\\[-0.5ex]
  \includegraphics[width=\textwidth]{../img/fitness_pfsp_rec19_txt}
\end{minipage}
\begin{minipage}{0.49\linewidth}
  \texttt{reC31} (PFSP)\\[-0.5ex]
  \includegraphics[width=\textwidth]{../img/fitness_pfsp_rec31_txt}
\end{minipage}
\label{fig:results}
\caption{Mean fitness  (and standard deviation)  of each solution evaluated by each algorithm on various instances.\label{fig:results}}
\end{figure*}
